---
title: "04_Stroud_Hannah_2018"
author: "Hannah Stroud"
date: "October 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hypothesis Testing 
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```


### 1- W&S, Ch 6, Question 15,21, and 29

####1.1 Question 15 
For the following alternative hypotheses, give the appropriate null hypothesis:  

15a. The null hypothesis would be: There are no differences in mean femur length between pygmy mammoths and continental mammoths.  

15b. Null: Patients who take phentermine and topiramate lose wait at the same rate as control patients that do not take these drugs.  

15c. Null:  Patients who take phentermine and topiramate have the same proportion of their babies born with cleft palates as patients who do not take these drugs.  

15d. Null: On average shoppers do not buy different amounts of candy when Christmas music is playing in the shop compared to when the usual type of music is playing.  

15e. Null: Male white-collared manakins dance as frequently when females are present as they do when females are absent.  

####1.2 Question 21
Two researchers testing the same null hypothesis, all equal example n1=60 and n2=100  

21a. The study with the higher probability of Type II error is the one with 60 participants because the study with 100 has a higher power, and increased power decreases the chances of a type II error.

21b. The study with the higher power is the one with 100 participants since all else is equal in the study- power increases with sample size 

21c. Type I errors are regulated by the choice of alpha, which is the same for both studies, so they are both as likely to commit a type I error....

21d. The test should be two-tailed because the null hypothesis is that the inhibitors have no effect on cardiac arrest, which means the alternative hypothesis is that they have an effect which could either be in either direction- reducing or increasing the risk of cardiac arrest.   

####1.3 Question 29
100 independent hypothesis tests with an alpha of 0.05. Use R to calc P value.

29a. If all null hypothesis were true, what is the probability that the researchers would reject none of them?
```{r}

```
29b. If all 100 null hypothesis were true, how many of those tests on average are expected to reject the null? 

### Question 2- ch 7, Q22

9821 slices of toast dropped, 6101 landed butter side down

a. What is the 95% CI for the probability of a toast landing butter side down?

b. Using the results of part a, is it plausible there is a 50-50 chance of toast landing butter size up or butter size down?  


###3 From the Lab: Many SDs and Alphas
Assume average population-wide resting heart rate of 80 bpm with a standard deviation of 6, effect size of 5  
#### Goal: what sample size do we need to achieve a power of 0.8? 

#### 3.1 Start up your simulation
Multiple sample sizes: from 1-20, with 500 simulations per sample size  
multiple SD values, from 3 through 10 (just 3:10, no need for non-integer values)  
You’re going to want crossing with your intitial data frame of just sample sizes and a vector of sd values to start. Then generate samples from the appropriate random normal distribution.

```{r}
Sim_data <- data.frame(samp_size = rep(1:20, 500))
SD_val <- c(3:10)
Sim_data <- crossing(Sim_data
         , SD_val)
Sim_data <- Sim_data %>% 
  
 group_by(1:n()) %>%   #n= allows us to change the 500, and not need to change 
  #get a sample mean
  mutate(sample_mean= mean(rnorm(samp_size, 85, SD_val))) %>%
  ungroup()
```

#### 3.2 Z Test
now that you've done that, calculate the results from z-tests. Plot p by sample size, using `facet_wrap` for different SD values  

```{r}
Sim_data <- Sim_data %>% 
 mutate(sigma_ybar= SD_val/(sqrt(samp_size))) %>%  #Sigma_y bar formula from slides
  mutate(Z_score= (sample_mean - 80)/sigma_ybar) %>%
  mutate(p_val= pnorm(abs(Z_score), lower.tail = FALSE))%>% 
  ungroup()

Sim_plot <- ggplot(data= Sim_data, 
                   mapping = aes(x= samp_size, y= p_val))+ 
  geom_point() +
  facet_wrap(SD_val)
Sim_plot

```
#### 3.3 P and Power
plot power for an alpha of 0.05, but use color for different SD values. Include our threshold power of 0.8.

```{r}
sim_power <- Sim_data %>%
  group_by(SD_val, samp_size) %>%
  summarize(power= 1-sum(p_val>0.05)/n()) %>%
  ungroup()

ggplot(data= sim_power, 
       mapping= aes(x= samp_size, y= power, group= SD_val, color= factor(SD_val)))+
         geom_point() +
          geom_line() +
    geom_hline(yintercept= 0.8, lty= 2)+
  scale_color_discrete(guide = guide_legend(title=expression(alpha)))
```

#### 3.4 Many alphas
Last, use `crossing` again to explore changing alphas from 0.01 to 0.1. Plot power curves with different alphas as different colors, and use faceting to look at different SDs.

```{r}
sim_alpha <- Sim_data %>% 
  crossing(alpha=10^(-2:-1)) %>%
  group_by(samp_size, SD_val, alpha) %>%
  summarize(power= 1- sum(p_val>alpha)/n()) %>%
  ungroup()

ggplot(data=sim_alpha, mapping= 
         aes(x= samp_size, y= power, color= factor(alpha)))+
  geom_point()+
  facet_wrap(~SD_val)


```

#### 3.5 What does it all mean? What do you learn about how alpha and SD affect power?  
The higher the standard deviation, the large sample size needed to get a satisfactory power. Increasing alpha increases power, but it leads to trade off of increasing your likelihood of falsely rejecting the null hypothesis.   

#### 3.6 How do you think that changing the effect size would affect power? 
A bigger effect size would increase power, because it would be easier to detect. 